---
title: "Weather and Corn Yield Regressions"
author: "Brandon Budnicki"
date: "3-3-2022"
output: html_document

knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_file = 'index',
      output_dir='./'
    )
  })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(R.matlab)
library(rnassqs)
library(gridExtra)
library(mapview)
library(USAboundaries)
library(USAboundariesData)
library(leaflet.extras2)
library(raster)


```

# Setup
## Load the PRISM daily maximum temperatures

```{r tmax data}

# daily max temperature
# dimensions: counties x days x years
prism <- readMat("prismiowa.mat")

# look at county #1
t_1981_c1 <- prism$tmaxdaily.iowa[1,,1]
t_1981_c1[366]
plot(1:366, t_1981_c1, type = "l")

ggplot() +
  geom_line(mapping = aes(x=1:366, y = t_1981_c1)) +
  theme_bw() +
  xlab("day of year") +
  ylab("daily maximum temperature (°C)") +
  ggtitle("Daily Maximum Temperature, Iowa County #1")


```
```{r tidying up}

# assign dimension names to tmax matrix
dimnames(prism$tmaxdaily.iowa) <- list(prism$COUNTYFP, 1:366, prism$years)

# converted 3d matrix into a data frame
tmaxdf <- as.data.frame.table(prism$tmaxdaily.iowa)

# relabel the columns
colnames(tmaxdf) <- c("countyfp","doy","year","tmax")
tmaxdf <- tibble(tmaxdf)

```

## Temperature trends

### Summer temperature trends: Winneshiek County

```{r temp trends}

tmaxdf$doy <- as.numeric(tmaxdf$doy)
tmaxdf$year <- as.numeric(as.character(tmaxdf$year))

winneSummer <- tmaxdf %>%
  filter(countyfp==191 & doy >= 152 & doy <= 243) %>%
  group_by(year) %>%
  summarize(meanTmax = mean(tmax))

ggplot(winneSummer, mapping = aes(x = year, y = meanTmax)) +
  geom_point() +
  theme_bw() +
  labs(x = "year", y = "Tmax (°C)") +
  geom_smooth(method = lm)

lm_summerTmax <- lm(meanTmax ~ year, winneSummer)
summary(lm_summerTmax)

```

### Winter Temperatures - Winneshiek County

```{r winter temps}

winneWinter <- tmaxdf %>%
  filter(countyfp==191 & (doy <= 59 | doy >= 335) & !is.na(tmax)) %>%
  group_by(year) %>%
  summarize(meanTmax = mean(tmax))

ggplot(winneWinter, mapping = aes(x = year, y = meanTmax)) +
  geom_point() +
  theme_bw() +
  labs(x = "year", y = "Tmax (°C)") +
  geom_smooth(method = lm)

lm_winterTmax <- lm(meanTmax ~ year, winneWinter)
summary(lm_winterTmax)

```

### Multiple regression -- Quadratic time trend

```{r quadratic temp trend}

winneWinter$yearSq <- winneWinter$year^2

lm_winterTmaxQuad <- lm(meanTmax ~ year + yearSq, winneWinter)
summary(lm_winterTmaxQuad)
winneWinter$fitted <- lm_winterTmaxQuad$fitted.values

ggplot(winneWinter) +
  geom_point(mapping = aes(x = year, y = meanTmax)) +
  geom_line(mapping = aes(x = year, y = fitted)) +
  theme_bw() +
  labs(x = "year", y = "Tmax (°C)")

```

### Download NASS corn yield data

```{r yield download, include=FALSE}

# set our API key with NASS
nassqs_key <- Sys.getenv("nassqs_auth")
nassqs_auth(key = nassqs_key)

# parameters to query on 
params <- list(commodity_desc = "CORN", util_practice_desc = "GRAIN", prodn_practice_desc = "ALL PRODUCTION PRACTICES", year__GE = 1981, state_alpha = "IA")

# download
cornyieldsall <- nassqs_yields(params)

cornyieldsall$county_ansi <- as.numeric(cornyieldsall$county_ansi)
cornyieldsall$yield <- as.numeric(cornyieldsall$Value)

# clean and filter this dataset
cornyields <- dplyr::select(cornyieldsall, county_ansi, county_name, yield, year) %>%
  filter(!is.na(county_ansi) & !is.na(yield))
cornyields <- tibble(cornyields)

```

# Assignment

## Question 1

### Question 1a

Extract Winneshiek County corn yields, fit a linear time trend, make a plot. Is there a significant time trend?

```{r, fig.width = 8, fig.height= 8, res=600}


winneYield = cornyields %>%
  filter(county_ansi==191)

ggplot(winneYield, mapping = aes(x = year, y = yield)) +
  geom_point( alpha = 0.4, colour = "black") +
  theme_bw() +
  labs(x = "Year", y = "Yield (bsh/ac)", title= "Yield over time", subtitle="Winneshiek") +
  geom_smooth(method = lm)

```

Corn yield has consistently increased year over year. This agrees with my prior knowledge from working at Land O'Lakes. Improvements in farming practices and seed genetics have been driving this trend. Although adoption of these changes varies from farm to farm, at the aggregated county level this affect can be clearly seen in the data.


### Question 1b

Fit a quadratic time trend (i.e., year + year^2) and make a plot. Is there evidence for slowing yield growth? 

```{r, fig.width = 9, fig.height= 8, res=600}
winneYieldLin<- lm(yield ~ year, winneYield)
summary(winneYieldLin)
winneYield$lin <- winneYieldLin$fitted.values


winneYield$yearSq <- winneYield$year^2
winneYieldQuad<- lm(yield ~ year + yearSq, winneYield)
summary(winneYieldQuad)
winneYield$quad <- winneYieldQuad$fitted.values

ggplot(winneYield) +
  geom_point(mapping = aes(x = year, y = yield), alpha = 0.4, colour = "black") +
  geom_line(mapping = aes(x = year, y = quad, col='Quadratic Time Model'), size=1) +
  geom_line(mapping = aes(x = year, y = lin, col='Linear Time Model'), size=1) +
  theme_bw() +
  labs(x = "Year", y = "Yield (bsh/ac)", title= "Yield over time", subtitle="Winneshiek, IA")

```

There is no evidence of slowing corn yield growth in Winneshiek County, IA. The slope of the quadratic time model is relatively constant over time.


## Question 2 -- Time Series

Let's analyze the relationship between temperature and yields for the Winneshiek County time series. Use data on **yield** and **summer avg Tmax**. Is adding year or Tmax^2 to your model helpful? Make a plot and interpret the results.

```{r}
winneYieldTemp = merge(winneYield, winneSummer, by = 'year') %>%
  mutate(
    yieldSq = yield^2,
    meanTmaxSq = meanTmax^2
  )

```
```{r, fig.width = 8, fig.height = 5}
winneYieldTempLin<- lm(yield ~ meanTmax, winneYieldTemp)
summary(winneYieldTempLin)
winneYieldTemp$tempLin <- winneYieldTempLin$fitted.values

winneYieldTempQuad<- lm(yield ~ meanTmax + meanTmaxSq, winneYieldTemp)
winneYieldTemp$tempQuad <- winneYieldTempQuad$fitted.values

winneYieldYearLin<- lm(yield ~ meanTmax + year, winneYieldTemp)
winneYieldTemp$yearLin <- winneYieldYearLin$fitted.values

winneYieldYearTempQuad<- lm(yield ~ year + yearSq, winneYieldTemp)
winneYieldTemp$yearTempQuad <- winneYieldYearTempQuad$fitted.values

```
```{r, fig.width = 8, fig.height= 8 , res=600}
grid.arrange(

ggplot(winneYieldTemp) +
  geom_point(mapping = aes(x = meanTmax, y = yield), alpha = 1, colour = "black") +
  geom_line(mapping = aes(x = meanTmax, y = tempQuad, col="Quadratic Temp Model"), size=1) +
  geom_line(mapping = aes(x = meanTmax, y = tempLin, col="Linear Temp Model"), size=1) +
  geom_line(mapping = aes(x = meanTmax, y = yearLin, col="Linear Temp Year Model"), size=1) +
  geom_line(mapping = aes(x = meanTmax, y = yearTempQuad, col="Quadratic Temp Year Model"), size=1) +
  theme_bw() +
  labs(x = "T Max (°C)", y = "Yield (bsh/ac)", title= "Yield over T Max", subtitle="Winneshiek, IA")

,

ggplot(winneYieldTemp) +
  geom_point(mapping = aes(x = year, y = yield), alpha = 1, colour = "black") +
  geom_line(mapping = aes(x = year, y = tempQuad, col="Quadratic Temp Model"), size=1) +
  geom_line(mapping = aes(x = year, y = tempLin, col="Linear Temp Model"), size=1) +
  geom_line(mapping = aes(x = year, y = yearLin, col="Linear Temp Year Model"), size=1) +
  geom_line(mapping = aes(x = year, y = yearTempQuad, col="Quadratic Temp Year Model"), size=1) +
  theme_bw() +
  labs(x = "Year", y = "Yield (bsh/ac)", title= "Yield over time", subtitle="Winneshiek, IA")

)
```

The yield peaks at a mean summer temperature max of 27 Celsius.This follows intuition that plants respond well to increasing temperature, if it was freezing they would not grow. However, increased temperature only increases yield up to 27 degrees than than gradually drops. The quadratic temperature model has significantly greater predictive performance than the linear. However with such a significant year over year affect, the Quadratic Temp Year model has the best performance.


## Question 3 -- Cross-Section

Analyze the relationship between temperature and yield across all counties in 2018. Is there a relationship? Interpret the results.

```{r}
tmaxAll = tmaxdf %>%
  filter(doy >= 152 & doy <= 243) %>%
  group_by(year, countyfp) %>%
  summarize(meanTmax = mean(tmax))

iowa = merge(cornyields, tmaxAll, by.x=c("year", "county_ansi"), by.y=c("year", "countyfp")) %>% 
  mutate(
    yearSq = year^2,
    yieldSq = yield^2,
    meanTmaxSq = meanTmax ^ 2
  )

iowa2018 = iowa %>%
  filter(year == 2018)

```


```{r, fig.width = 8, fig.height= 8, res=600}
iowa2018YieldTempLin<- lm(yield ~ meanTmax, iowa2018)
summary(iowa2018YieldTempLin)
iowa2018$lin <- iowa2018YieldTempLin$fitted.values

iowa2018YieldTempQuad<- lm(yield ~ meanTmax + meanTmaxSq, iowa2018)
summary(iowa2018YieldTempQuad)
iowa2018$quad <- iowa2018YieldTempQuad$fitted.values

ggplot(iowa2018) +
  geom_point(mapping = aes(x = meanTmax, y = yield), alpha = 0.4, colour = "black") +
  geom_line(mapping = aes(x = meanTmax, y = quad, colour="Quadratic"), size=2) +
  geom_line(mapping = aes(x = meanTmax, y = lin, colour="Linear"), size=2) +
  theme_bw() +
  labs(x = "T Max (°C)", y = "Yield (bsh/ac)", title= "Yield over T Max", subtitle="Iowa 2018")


```

We see the same relationship between Summer Mean T Max and Yield in 2018 for all counties in Iowa that we saw for all years in Winneshiek. The peak is at 28 degrees instead of 27 degrees. This data tells the story better, as unlike the Winneshiek analysis, this is not affected by year of year yield increases due to improved farming methods.

## Question 4 -- Panel

One way to leverage multiple time series is to group all data into what is called a "panel" regression. Convert the county ID code ("countyfp" or "county_ansi") into factor using as.factor, then ***include this variable in a regression using all counties' yield and summer temperature data***. How does the significance of your temperature coefficients (Tmax, Tmax^2) change? Make a plot comparing actual and fitted yields and interpret the results of your model.

```{r}
# add factors
iowa$county_name = as.factor(iowa$county_name)
iowa$county_ansi = as.factor(iowa$county_ansi)

```
```{r}
# Single Factor Models
iowa_TempQuad = lm(yield ~ meanTmax + meanTmaxSq, iowa)
#summary(iowa_TempQuad)
iowa$m_tempSq <- iowa_TempQuad$fitted.values

iowa_YearQuad = lm(yield ~ year + yearSq, iowa)
#summary(iowa_YearQuad)
iowa$m_yearSq <- iowa_YearQuad$fitted.values

iowa_County = lm(yield ~ county_name, iowa)
#summary(iowa_County)
iowa$m_county <- iowa_County$fitted.values

iowa_Combined = lm(yield ~ year + yearSq + meanTmax + meanTmaxSq + county_name, iowa)
#summary(iowa_Combined)
iowa$m_combined <- iowa_Combined$fitted.values

```


```{r, fig.width = 8, fig.height= 12, res=600}
grid.arrange(
top = "Single Factor Models",
layout_matrix = rbind(c(1, 2, 3),c(4, 4, 4)),

ggplot(iowa) +
  geom_point(mapping = aes(x = yield, y = m_tempSq), alpha = 0.2) +
  geom_line(mapping = aes(x = yield, y = yield), linetype = "dashed", size=1.5) +
  theme_bw() +
  labs(x = "Actual Yield", y = "Fitted Yield", title= "Quadratic Temp Model VS Actual",)
,
ggplot(iowa) +
  geom_point(mapping = aes(x = yield, y = m_yearSq), alpha = 0.2) + 
  geom_line(mapping = aes(x = yield, y = yield), linetype = "dashed", size=1.5) +
  theme_bw() +
  labs(x = "Actual Yield", y = "Fitted Yield", title= "Quadratic Year Model VS Actual")
,
ggplot(iowa) +
  geom_point(mapping = aes(x = yield, y = m_county), alpha = 0.2) + 
  geom_line(mapping = aes(x = yield, y = yield), linetype = "dashed", size=1.5) +
  theme_bw() +
  labs(x = "Actual Yield", y = "Fitted Yield", title= "County Model VS Actual")
,
ggplot(iowa) +
  geom_point(mapping = aes(x = yield, y = m_combined), alpha = 0.2) + 
  geom_line(mapping = aes(x = yield, y = yield), linetype = "dashed", size=1.5) +
  theme_bw() +
  labs(x = "Actual Yield", y = "Fitted Yield", title= "Combined Model VS Actual")

)


```

Combined the models tell a more complete picture. On its own, the quadratic temperture model only have predictive power up to 150 bushels, after which all increases are dependent on factors outside of temperature. The year model provides the best single factor fit, following the slope most tightly The county also plays a large factor in that a 50 bushel difference can be attributed to the location of your farm.

```{r}
# Progessive models
iowa_TempQuadModel = lm(yield ~ meanTmax + meanTmaxSq, iowa)
#summary(iowa_TempQuadModel)
iowa$tempQuad <- iowa_TempQuadModel$fitted.values

iowa_TempQuadCountyModel = lm(yield ~ meanTmax + meanTmaxSq + county_name, iowa)
#summary(iowa_TempQuadCountyModel)
iowa$tempQuadCounty <- iowa_TempQuadCountyModel$fitted.values

iowa_TempQuadYearQuadCountyLinModel = lm(yield ~ meanTmax + meanTmaxSq + year + yearSq + county_name, iowa)
#summary(iowa_TempQuadYearQuadCountyLinModel)
iowa$tempQuadYearQuadCountyLin <- iowa_TempQuadYearQuadCountyLinModel$fitted.values



```

```{r, fig.width = 10, fig.height= 8}

# ggplot(iowa) +
#   geom_point(mapping = aes(x = meanTmax, y = yield), alpha = 0.4, colour = "black") +
#   geom_line(mapping = aes(x = meanTmax, y = tempQuadYearQuadCountyQuad, col="Quadratic Temp, Quadratic Year, Linear County Model")) + 
#   geom_line(mapping = aes(x = meanTmax, y = tempQuadCounty, col="Quadratic Temp, Linear County Model")) + 
#   geom_line(mapping = aes(x = meanTmax, y = tempQuad, col="Quadratic Temp Model")) +
#   theme_bw() +
#   labs(x = "T Max (°C)", y = "Yield (bsh/ac)", title= "Yield over T Max", subtitle="Iowa All Years")
# 
# 
# 
# ggplot(iowa) +
#   geom_point(mapping = aes(x = year, y = yield), alpha = 0.4, colour = "black") +
#   geom_line(mapping = aes(x = year, y = tempQuadYearQuadCountyQuad, col="Quadratic Temp, Quadratic Year, Linear County Model")) + 
#   geom_line(mapping = aes(x = year, y = tempQuadCounty, col="Quadratic Temp, Linear County Model")) + 
#   geom_line(mapping = aes(x = year, y = tempQuad, col="Quadratic Temp Model")) +
#   theme_bw() +
#   labs(x = "Year", y = "Yield (bsh/ac)", title= "Yield over time", subtitle="Iowa All Years")
# 
# 
# 
# 
# ggplot(iowa) +
#   geom_point(mapping = aes(x = county_name, y = yield), alpha = 0.4, colour = "black") +
#   geom_line(mapping = aes(x = county_name, y = tempQuadYearQuadCountyQuad, col="Quadratic Temp, Quadratic Year, Linear County Model")) + 
#   geom_line(mapping = aes(x = county_name, y = tempQuadCounty, col="Quadratic Temp, Linear County Model")) + 
#   geom_line(mapping = aes(x = county_name, y = tempQuad, col="Quadratic Temp Model")) +
#   theme_bw() +
#   labs(x = "Year", y = "Yield (bsh/ac)", title= "Yield over County", subtitle="Iowa All Years")
# 

```
```{r, fig.width = 10, fig.height= 16, res=600}
grid.arrange(
top = "Progressively more complex models",

ggplot(iowa) +
  geom_point(mapping = aes(x = yield, y = tempQuad), alpha = 0.2) +
  geom_line(mapping = aes(x = yield, y = yield), linetype = "dashed", size=2) +
  theme_bw() +
  labs(x = "Actual Yield", y = "Fitted Yield", title= "Quadratic Temp Model VS Actual", subtitle="Iowa All Years")
,
ggplot(iowa) +
  geom_point(mapping = aes(x = yield, y = tempQuadCounty), alpha = 0.2) + 
  geom_line(mapping = aes(x = yield, y = yield), linetype = "dashed", size=2) +

    theme_bw() +
  labs(x = "Actual Yield", y = "Fitted Yield", title= "Quadratic Temp, County Model VS Actual", subtitle="Iowa All Years")
,
ggplot(iowa) +
  geom_point(mapping = aes(x = yield, y = tempQuadYearQuadCountyLin), alpha = 0.2) + 
  geom_line(mapping = aes(x = yield, y = yield), linetype = "dashed", size=2) +

    theme_bw() +
  labs(x = "Actual Yield", y = "Fitted Yield", title= "Quadratic Temp, Quadratic Year, County Model VS Actual", subtitle="Iowa All Years")

)
```

## Question 5 -- Soybeans

Download NASS data on soybean yields and explore either a time series relationship for a given county, the cross-sectional relationship for a given year, or a panel across all counties and years.

```{r, include=FALSE}
## download, format, & clean data
soyParams <- list(
  commodity_desc = "SOYBEANs", statisticcat_desc="YIELD", 
  prodn_practice_desc = "ALL PRODUCTION PRACTICES", year__GE = 1981, state_alpha = "IA")

soyYields<- nassqs_yields(soyParams)

soyYields$county_ansi <- as.numeric(soyYields$county_ansi)
soyYields$yield <- as.numeric(soyYields$Value)

# clean and filter this dataset
soyYields <- dplyr::select(soyYields, county_ansi, county_name, yield, year) %>%
  filter(!is.na(county_ansi) & !is.na(yield))


soy = merge(soyYields, tmaxAll, by.x=c("year", "county_ansi"), by.y=c("year", "countyfp")) %>% 
  mutate(
    yearSq = year^2,
    yieldSq = yield^2,
    meanTmaxSq = meanTmax ^ 2,
    county_name = as.factor(tolower(county_name)),
    county_ansi = as.factor(county_ansi)
  )
```

```{r}
soyMeanTmaxQuadModel<- lm(yield ~ meanTmax + meanTmaxSq, soy)
soy$m_tmax_quad <- soyMeanTmaxQuadModel$fitted.values
soyMeanTmaxLinModel<- lm(yield ~ meanTmax, soy)
soy$m_tmax_lin <- soyMeanTmaxLinModel$fitted.values

soyYearQuadModel<- lm(yield ~ year + yearSq, soy)
soy$m_year_quad <- soyYearQuadModel$fitted.values
soyYearLinModel<- lm(yield ~ year, soy)
soy$m_year_lin <- soyYearLinModel$fitted.values

soyCountyModel<- lm(yield ~ county_name, soy)
soy$m_county <- soyCountyModel$fitted.values


soyCombinedModel<- lm(yield ~ meanTmax + meanTmaxSq + year + yearSq + county_name, soy)
soy$m_combined <- soyCombinedModel$fitted.values

```

```{r}
plot_temp = ggplot(soy) +
  geom_point(mapping = aes(x = meanTmax, y = yield), alpha = 0.2, colour = "black") +
  geom_line(mapping = aes(x = meanTmax, y = m_tmax_quad, colour="Quadratic"), size=2) +
  geom_line(mapping = aes(x = meanTmax, y = m_tmax_lin, colour="Linear"), size=2) +
  theme_bw() +
  labs(x = "T Max (°C)", y = "Yield (bsh/ac)", title= "Yield over T Max")

plot_year = ggplot(soy) +
  geom_point(mapping = aes(x = year, y = yield), alpha = 0.2, colour = "black") +
  geom_line(mapping = aes(x = year, y = m_year_quad, colour="Quadratic"), size=2) +
  geom_line(mapping = aes(x = year, y = m_year_lin, colour="Linear"), size=2) +
  theme_bw() +
  labs(x = "Year", y = "Yield (bsh/ac)", title= "Yield over Year")

plot_county = ggplot(soy) +
  geom_point(mapping = aes(x = county_name, y = yield), alpha = 0.2, colour = "black") +
  geom_line(mapping = aes(x = county_name, y = m_county), size=2) +
  theme_bw() +
  labs(x = "County", y = "Yield (bsh/ac)", title= "Yield over County")

```
```{r}

plot_model_temp = ggplot(soy) +
  geom_point(mapping = aes(x = yield, y = m_tmax_quad), alpha = 0.2) +
  geom_line(mapping = aes(x = yield, y = yield), linetype = "dashed", size=2) +
  theme_bw() +
  labs(x = "Actual Yield", y = "Fitted Yield", title= "Quadratic Temp Model", subtitle="Iowa All Years")

plot_model_year = ggplot(soy) +
  geom_point(mapping = aes(x = yield, y = m_year_quad), alpha = 0.2) + 
  geom_line(mapping = aes(x = yield, y = yield), linetype = "dashed", size=2) +
  theme_bw() +
  labs(x = "Actual Yield", y = "Fitted Yield", title= "Quadratic Year Model", subtitle="Iowa All Years")

plot_model_county = ggplot(soy) +
  geom_point(mapping = aes(x = yield, y = m_county), alpha = 0.2) + 
  geom_line(mapping = aes(x = yield, y = yield), linetype = "dashed", size=2) +
  theme_bw() +
  labs(x = "Actual Yield", y = "Fitted Yield", title= "County Model", subtitle="Iowa All Years")

```

```{r}
plot_model_combined = ggplot(soy) +
  geom_point(mapping = aes(x = yield, y = m_combined), alpha = 0.2, size=2) + 
  geom_line(mapping = aes(x = yield, y = yield), linetype = "dashed") +
  theme_bw() +
  labs(x = "Actual Yield", y = "Fitted Yield", title= "Combined Model VS Actual", subtitle="Iowa All Years")

```

```{r,fig.width = 10, fig.height= 8, res=600}


grid.arrange(
top = "Factors of Soybean Yield in Iowa",
layout_matrix = rbind(
  c(1,2,3),
  c(4,5,6),
  c(7)
),
plot_temp, plot_year, plot_county,
plot_model_temp, plot_model_year, plot_model_county,
plot_model_combined
)


```



## Question 6 Bonus

Find a package to make a county map of Iowa displaying some sort of information about yields or weather. Interpret your map.

```{r}
counties = us_counties(map_date = NULL, resolution = c("low", "high"), states = 'iowa') %>%
  dplyr::select(name, geometry)
counties$name = tolower(counties$name)

```
```{r fig.width = 10, fig.height= 8}
selectYear = 2018

cornTime = cornyields %>% 
  filter(year == selectYear) %>%
  mutate(
    yearSq = year^2,
    yieldSq = yield^2,
    county_name = as.factor(tolower(county_name)),
    county_ansi = as.factor(county_ansi)
  )

soyTime = soy %>% 
  filter(year == selectYear)


mapCorn = merge(counties, cornTime, by.y = "county_name", by.x = "name", all.x=TRUE)
mapSoy = merge(counties, soyTime, by.y = "county_name", by.x = "name", all.x=TRUE)

mapview(mapCorn, alpha.regions = 0.5, aplha = 1, zcol="yield", 
  legend = TRUE,
  na.color = '#777777'
) |

mapview(mapSoy, alpha.regions = 0.5, aplha = 1, zcol="yield",
  legend = TRUE,
  na.color = '#777777'

)


```

## Question 7 Bonus 

Map trends in corn yields by county across Iowa. Interpret your map.

```{r fig.width = 10, fig.height= 8}
brks = classInt::classIntervals(
  1:100,
  style = 'fixed',
  fixedBreaks = seq(24,68,6),
  intervalClosure = "right"
)


soyEarly = soy %>% 
  filter(year == 1985)
soyLate = soy %>% 
  filter(year == 2017) 

mapSoyEarly = merge(counties, soyEarly, by.y = "county_name", by.x = "name", all.x=TRUE)
mapSoyLate = merge(counties, soyLate, by.y = "county_name", by.x = "name", all.x=TRUE)

mapview(mapSoyEarly, alpha.regions = 0.5, aplha = 1, zcol="yield", 
  at = brks$brks,
  legend = FALSE,
  na.color = '#777777'
) |

mapview(mapSoyLate, alpha.regions = 0.5, aplha = 1, zcol="yield",
  at = brks$brks,
    na.color = '#777777'

)


```



# Links

[Git Hub Repo](https://github.com/BBudnicki/weather_corn_regressions)

[Git Hub Pages](https://bbudnicki.github.io/weather_corn_regressions/)
